---
title: "p8105_hw5_enp2116"
author: "Emily Potts"
date: "2022-11-13"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(gridExtra)

set.seed(4)
```

# Problem 1
* Starting with a dataframe containing all file names, iterating over file names and reading in data for each subject, using purrr::map and saving the result as a new variable in the dataframe:
```{r, message = FALSE}
all_files_df = 
  tibble(
    files = list.files("data/zip_data/"),
    path = str_c("data/zip_data/", files)
  ) %>% 
  mutate(data = purrr::map(path, read_csv)) %>% 
  unnest(cols = c(data))
head(all_files_df)
```

* Tidying the dataframe: manipulating file names to include control arm and subject ID, make sure weekly observations are tidy, and making sure variables are the correct type.
```{r}
tidy_df = 
  all_files_df %>% 
  mutate(
    files = str_replace(files, ".csv", ""),
    group = str_sub(files, 1, 3)) %>% 
  pivot_longer(
    week_1:week_8,
    names_to = "week",
    values_to = "outcome",
    names_prefix = "week_") %>% 
  mutate(week = as.numeric(week)) %>% 
  select(group, subj = files, week, outcome)
head(tidy_df)
```

* Making a spaghetti plot using ggplot to show observations on each subject over time:
```{r, echo = FALSE}
tidy_df %>% 
  ggplot(aes(x = week, y = outcome, group = subj, color = group)) + 
  geom_point() + 
  geom_path() + xlab("Week") + ylab("Outcome") + 
  labs(title = "Observations Over Time by Subject") + facet_grid(~group)
```

* Differences between groups:

The side-by-side spaghetti plots show differences between the control and experimental groups. While both groups start off around the same level of outcome (experimental slightly higher), the control group is stable over time. In contrast, the experimental group increases relatively linearly over the time of the study. Finally, we see high within-subject correlation, where subjects who start above average end up above average, and those that start below average end up below average. 

# Problem 2

```{r, message = FALSE, echo = FALSE}
raw_homicide_df =
  read_csv("data/homicide-data.csv")
```
* The raw homicide data includes information on the victims (first name, last name, race, age, sex), information on the report (ID, date reported), information on the location of the homicide (city, state, latitude, longitude), and the outcome of the investigation (disposition). There are `r nrow(raw_homicide_df)` rows, which means we have data on `r nrow(raw_homicide_df)` different homicides across the United States. We have a variable of interest disposition, which takes on the values of`r unique(raw_homicide_df$disposition)`.

* Creating a city_state variable and summarizing within cities to obtain the total number of homicides and the number of unsolved homicides:
```{r}
tidy_homicide = 
  raw_homicide_df %>% 
  mutate(
    city_state = paste(city, state, sep = ", "),
    city_state = ifelse(city_state == "Tulsa, AL", "Tulsa, OK", city_state),
    outcome = case_when(disposition == "Closed without arrest" ~ 'unsolved',
                        disposition == "Open/No arrest" ~ 'unsolved',
                        disposition == "Closed by arrest" ~ 'solved')) %>% 
  group_by(city_state) %>% 
  summarize(
    unsolved_n = sum(outcome == "unsolved"),
    total_n = n()
  ) 
head(tidy_homicide)
```

* For the city of Baltimore, MD, using the prop.test function to estimate the proportion of homicides that are unsolved; saving the output of prop.test as an R object, applying the broom::tidy to this object and pulling the estimated proportion and confidence intervals from the resulting tidy dataframe:
```{r}
x = tidy_homicide %>% 
  filter(city_state == "Baltimore, MD") %>% 
  pull(unsolved_n)

n = tidy_homicide %>% 
  filter(city_state == "Baltimore, MD") %>% 
  pull(total_n)

prop.test(x, n) %>% 
  broom::tidy() %>% 
  select(estimate, starts_with("conf"))
```

* Running prop.test for each of the cities in your dataset, and extracting both the proportion of unsolved homicides and the confidence interval for each:
```{r}
homicide_prop_tests = 
  tidy_homicide %>% 
  group_by(city_state) %>% 
  mutate(
    h_prop_test = map2(unsolved_n, total_n, ~prop.test(.x, .y) %>% broom::tidy())) %>% 
      unnest(h_prop_test) %>% 
      select(city_state, estimate, conf.low, conf.high)
head(homicide_prop_tests)  
```

* Creating a plot that shows the estimates and CIs for each city using geom_errorbar for a way to add error bars based on the upper and lower limits
```{r, echo = FALSE}
homicide_prop_tests %>% 
  ggplot(aes(x = fct_reorder(city_state, estimate), y = estimate, color = city_state)) + geom_point() +
  geom_errorbar(aes(ymin = conf.low, ymax = conf.high)) + theme(axis.text.x = element_text(angle = 90), legend.position = "none") + labs(x = "Location of Homicide (City, State)", y = "Estimated Proportion of Unsolved Homicides", title = "95% CI of Proportion Unsolved Homicides by City")
```


# Problem 3

Fixing $n = 30, \sigma = 5$, setting $\mu = 0$ and generating 5000 datasets from the model. For each dataset, we save $\hat{\mu}$ and the p-value arising from a test of if $\mu = 0$ for $\alpha = 0.05$. 
```{r, cache = TRUE}
sim_mu = function(n_obs = 30, mu, sigma = 5) {
  
  sim_data = tibble(
    x = rnorm(n = n_obs, mu, sd = sigma),
  )
  
  sim_data %>% 
    summarize(
      mu_hat = mean(x),
      p_value = (broom::tidy(t.test(x, mu = 0, conf.level = 0.95))$p.value)
    )
}

sim_mu_zero = 
  expand_grid(
    mu_value = 0,
    iter = 1:5000
  ) %>% 
  mutate(
    estimate_df = map(.x = mu_value, ~ sim_mu(mu = .x))
  ) %>% 
  unnest(estimate_df)
head(sim_mu_zero)
```

Repeating the above for $\mu={1,2,3,4,5,6}$:
```{r, cache = TRUE}
sim_results_df = 
  expand_grid(
    mu_list = c(1,2,3,4,5,6),
    iter = 1:5000
  ) %>% 
  mutate(
    estimate_df = map(.x = mu_list, ~ sim_mu(mu = .x))
  ) %>% 
  unnest(estimate_df)
sim_results_df
```

* Make a plot showing the proportion of times the null was rejected (the power of the test) on the y axis and the true value of $\mu$ on the x axis. Describe the association between effect size and power.
```{r, echo=FALSE}
sim_results_df %>% 
  mutate(
    reject = case_when(p_value <= 0.05  ~ TRUE,
                        p_value > 0.05 ~ FALSE)) %>% 
  group_by(mu_list) %>% 
  mutate(power = mean(reject)) %>%
  ggplot(aes(x = mu_list, y = power)) + geom_line() + geom_point() + xlab("True Population Mean") + ylab("Power") + labs(title = "True Population Mean vs Power")
```
From the plot of true population mean versus power, we can see that at a true $\mu$ of 1, the power of the test is quite low. However, as $\mu$ increases, the power increases at an decreasing rate as it approaches 1. Generalizing this relationship from our example, we can see that as effect size increases from the null (which is $\mu = 0$ here), we have greater power to detect a true difference and at a high enough effect size, the tests will be almost perfect at doing so.

* Making a plot showing the average estimate of $\hat{\mu}$ on the y axis and the true value of $\mu$ on the x axis. Second, making a second plot of the average estimate of $\hat{\mu}$ only in samples for which the null was rejected on the y axis and the true value of $\mu$ on the x axis. 
```{r, echo = FALSE, warning = FALSE}
plot_1 = sim_results_df %>% 
  group_by(mu_list) %>%
  mutate(
    reject = case_when(p_value <= 0.05  ~ TRUE,
                        p_value > 0.05 ~ FALSE),
    avg_mu_hat = mean(mu_hat)
    ) %>% 
  ggplot(aes(mu_list)) + geom_line(aes(y = avg_mu_hat), color = "blue") + geom_point(aes(y = avg_mu_hat), color = "blue") + xlab("True Population Mean") + ylab("Avg Estimate of Pop Mean") + labs(title = "True Population Mean vs Average Estimate") + ylim(0, 7)

plot_2 = sim_results_df %>% 
  group_by(mu_list) %>%
  mutate(
    reject = case_when(p_value <= 0.05  ~ TRUE,
                        p_value > 0.05 ~ FALSE)) %>% 
  filter(reject == TRUE) %>% 
    mutate(avg_mu_hat_reject = mean(mu_hat)) %>% 
  ggplot(aes(mu_list)) + geom_line(aes(y = avg_mu_hat_reject), color = "red") + geom_point(aes(y = avg_mu_hat_reject), color = "red") + xlab("True Population Mean") + ylab("Avg Estimate of Pop Mean") + labs(title = "True Population Mean vs Average Estimate for Rejected Null") + ylim(0, 7)

grid.arrange(plot_1, plot_2)
```
The sample average of $\hat{\mu}$ across tests for which the null is rejected approximately equal to the true value of $\mu$ at $\mu \geq 3$. At $\mu = 1$, we see a large discrepancy between the average estimates, with the rejected estimate over 2 and the total at about 1. The discrepancy gets smaller at $\mu = 2$, with the rejected estimate at about 2.5 and the total at about 2. This is because after the effect size gets large enough, the test will nearly always reject the null. At smaller values of $\mu$, where rejection occurs more often because values are closer to $\mu_0 = 0$, the rejected $\hat{\mu}$s are larger on average. 




