---
title: "p8105_hw5_enp2116"
author: "Emily Potts"
date: "2022-11-13"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(gridExtra)

set.seed(4)
```

# Problem 1
* Start with a dataframe containing all file names; the list.files function will help

* Iterate over file names and read in data for each subject using purrr::map and saving the result as a new variable in the dataframe

* Tidy the result; manipulate file names to include control arm and subject ID, make sure weekly observations are “tidy”, and do any other tidying that’s necessary

Make a spaghetti plot showing observations on each subject over time, and comment on differences between groups.


# Problem 2

* Describe the raw data. 

* Create a city_state variable (e.g. “Baltimore, MD”) and then summarize within cities to obtain the total number of homicides and the number of unsolved homicides (those for which the disposition is “Closed without arrest” or “Open/No arrest”).

* For the city of Baltimore, MD, use the prop.test function to estimate the proportion of homicides that are unsolved; save the output of prop.test as an R object, apply the broom::tidy to this object and pull the estimated proportion and confidence intervals from the resulting tidy dataframe.

* Now run prop.test for each of the cities in your dataset, and extract both the proportion of unsolved homicides and the confidence interval for each. Do this within a “tidy” pipeline, making use of purrr::map, purrr::map2, list columns and unnest as necessary to create a tidy dataframe with estimated proportions and CIs for each city.

Create a plot that shows the estimates and CIs for each city – check out geom_errorbar for a way to add error bars based on the upper and lower limits. Organize cities according to the proportion of unsolved homicides.




# Problem 3

Fixing $n = 30, \sigma = 5$, setting $\mu = 0$ and generating 5000 datasets from the model. For each dataset, we save $\hat{\mu}$ and the p-value arising from a test of if $\mu = 0$ for $\alpha = 0.05$. 
```{r}
sim_mean_sd = function(n_obs = 30, mu, sigma = 5) {
  
  sim_data = tibble(
    x = rnorm(n = n_obs, mu, sd = sigma),
  )
  
  sim_data %>% 
    summarize(
      mu_hat = mean(x),
      p_value = (broom::tidy(t.test(x, mu = 0, conf.level = 0.95))$p.value)
    )
}
```

Repeating the above for $\mu={1,2,3,4,5,6}$:
```{r}
sim_results_df = 
  expand_grid(
    mu_list = c(1,2,3,4,5,6),
    iter = 1:50
  ) %>% 
  mutate(
    estimate_df = map(.x = mu_list, ~ sim_mean_sd(mu = .x))
  ) %>% 
  unnest(estimate_df)
sim_results_df
```

* Make a plot showing the proportion of times the null was rejected (the power of the test) on the y axis and the true value of $\mu$ on the x axis. Describe the association between effect size and power.
```{r, echo=FALSE}
sim_results_df %>% 
  mutate(
    reject = case_when(p_value <= 0.05  ~ TRUE,
                        p_value > 0.05 ~ FALSE)) %>% 
  group_by(mu_list) %>% 
  mutate(power = mean(reject)) %>%
  ggplot(aes(x = mu_list, y = power)) + geom_line() + geom_point() + xlab("True Population Mean") + ylab("Power") + labs(title = "True Population Mean vs Power")
```
From the plot of true population mean versus power, we can see that at a true $\mu$ of 1, the power of the test is quite low. However, as $\mu$ increases, the power increases at an decreasing rate as it approaches 1. Generalizing this relationship from our example, we can see that as effect size increases from the null (which is $\mu = 0$ here), we have greater power to detect a true difference and at a high enough effect size, the tests will be almost perfect at doing so.

* First, making a plot showing the average estimate of $\hat{\mu}$ on the y axis and the true value of $\mu$ on the x axis. Second, making a second plot of the average estimate of $\hat{\mu}$ only in samples for which the null was rejected on the y axis and the true value of $\mu$ on the x axis. 
```{r, echo=FALSE}
plot_1 = sim_results_df %>% 
  group_by(mu_list) %>%
  mutate(
    reject = case_when(p_value <= 0.05  ~ TRUE,
                        p_value > 0.05 ~ FALSE),
    avg_mu_hat = mean(mu_hat)
    ) %>% 
  ggplot(aes(mu_list)) + geom_line(aes(y = avg_mu_hat), color = "blue") + geom_point(aes(y = avg_mu_hat), color = "blue") + xlab("True Population Mean") + ylab("Avg Estimate of Pop Mean") + labs(title = "True Population Mean vs Average Estimate") + ylim(0, 6)
```

```{r, echo=FALSE}
plot_2 = sim_results_df %>% 
  group_by(mu_list) %>%
  mutate(
    reject = case_when(p_value <= 0.05  ~ TRUE,
                        p_value > 0.05 ~ FALSE)) %>% 
  filter(reject == TRUE) %>% 
    mutate(avg_mu_hat_reject = mean(mu_hat)) %>% 
  ggplot(aes(mu_list)) + geom_line(aes(y = avg_mu_hat_reject), color = "red") + geom_point(aes(y = avg_mu_hat_reject), color = "red") + xlab("True Population Mean") + ylab("Avg Estimate of Pop Mean") + labs(title = "True Population Mean vs Average Estimate for Rejected Null") + ylim(0, 6)

grid.arrange(plot_1, plot_2)
```
The sample average of $\hat{\mu}$ across tests for which the null is rejected approximately equal to the true value of $\mu$ at $\mu \geq 3$. At $\mu = 1$, we see a large discrepancy between the average estimates, with the rejected estimate over 2 and the total at about 1. The discrepancy gets smaller at $\mu = 2$, with the rejected estimate at about 2.5 and the total at about 2. This is because after the effect size gets large enough, the test will nearly always reject the null. At smaller values of $\mu$, where rejection occurs more often because values are closer to $\mu_0 = 0$, the rejected $\hat{\mu}$s are larger on average.




